---
title: "Black Friday Analysis"
author: Shivangi
output: github_document
---
## Analysis of Black Friday Dataset

My work is to perform a preliminary analysis of consumer behavior using sales
data on **Black Friday**. I downloaded the
[BlackFriday](https://www.kaggle.com/llopesolivei/blackfriday#) dataset from Kaggle,
and loaded it into a new dataset called `blackfriday`.

```{r q1}
suppressPackageStartupMessages(library(tidyverse))
blackfriday <- read_csv("BlackFriday.csv", show_col_types = FALSE)
problems(blackfriday)
blackfriday %>% 
  ggplot(aes(x = Age)) + 
  geom_bar() + 
  labs(title = "Bar graph for Age")
blackfriday %>% 
  ggplot(aes(x = Marital_Status)) + 
  geom_bar() + 
  labs(title = "Bar graph for Marital Status")
blackfriday %>% 
  ggplot(aes(x = Purchase)) + 
  geom_density() + 
  labs(title = "Density Plot for Purchases")
```

As age and marital status are categorical variables, I am showing them in a bar graph. For purchases I am using a density graph. From the above three graphs, I can deduce that the data for Age, Marital Status and Purchases doesn't have any unusual values.

Here is a description of the variables.
"Masked" means that the values have been changed to be numeric values rather than the original categories. 

Variable                   | Definition
  ---                      |  ---
User_ID                    | User ID
Product_ID                 | Product ID
Gender                     | Sex of User
Age                        | Age in bins
Occupation                 | Occupation (Masked)
City_Category              | Category of the City (A,B,C)
Stay_In_Current_City_Years | Number of years stay in current city
Marital_Status             | Marital Status
Product_Category_1         | Product Category (Masked)
Product_Category_2         | Product may belongs to other category also (Masked)
Product_Category_3         | Product may belongs to other category also (Masked)
Purchase                   | Number of Purchases

```{r q2}
blackfriday <- unite(blackfriday, age_city, Age, City_Category)
blackfriday %>% 
  ggplot(aes(x = age_city, y = Purchase)) + 
  geom_boxplot() + 
  coord_flip()+
  labs(title = "Boxplot of Purchase by age_city")
```

Few age groups like 18-25, 26-35, 36-45 are making more purchases than others in especially city code C. For age group 55+, it is noticable that people in city code B are making more purchases than others in that age group. 

```{r q3}
blackfriday %>% filter(is.na(Product_Category_1))
blackfriday %>% filter(is.na(Product_Category_2), !is.na(Product_Category_3))
blackfriday_cat <- blackfriday %>% pivot_longer(c('Product_Category_1', 'Product_Category_2', 'Product_Category_3'), names_to = "Category_Order", values_to = "Prod_Category_ID", values_drop_na = TRUE)
blackfriday_cat
```

```{r q4}
blackfriday_cat %>% group_by(Prod_Category_ID) %>% mutate(distProdID = n_distinct(Product_ID)) %>% ungroup() %>% ggplot(aes(x = reorder(Prod_Category_ID, Purchase, sum), y = Purchase)) + geom_boxplot(aes(fill = distProdID)) + scale_fill_viridis_c() + labs(title = "Boxplot of Purchases by Product Category", x = "Product Category ID", fill = "Product ID")
```

We can notice that products with higher product IDs have been purchased lesser number of times than some of the products with lower number of purchases. 
The previous command to make product category ID as one single column instead of three columns helped visualise this data better. It was also beneficial for creating this boxplot. As it would have been a little difficult to visualize three different column in a single graph.